Integrating brain and biomechanical models - a new paradigm for understanding neuro-muscular control

### 2017-08-30 papers/integration/omm_integ.tex

%%% This is based on the Frontiers template.
\documentclass{frontiersSCNS}

@@

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{James {et~al.}}
\def\Authors{Sebastian James\,$^{1,2,*}$, Chris Papapavlou\,$^5$, Alexander Blenkinsop\,$^{1,2}$, \\
  Alexander Cope\,$^3$,  \\
  Sean Anderson\,$^{2,4}$, Konstantinos Moustakas\,$^5$, Kevin Gurney\,$^{1,2}$}
% Affiliations should be keyed to the author's name with superscript
% numbers and be listed as follows: Laboratory, Institute, Department,
% Organization, City, State abbreviation (USA, Canada, Australia), and
% Country (without detailed address information such as city zip codes
% or street names).
%
% If one of the authors has a change of address, list the new address
% below the correspondence details using a superscript symbol and use
% the same symbol to indicate the author in the author list.
\def\Address{\\
$^{1}$Adaptive Behaviour Research Group, Department of Psychology,
  The University of Sheffield, Sheffield, UK \\
$^{2}$Insigneo Institute for in-silico Medicine,
  The University of Sheffield, Sheffield, UK \\
$^{3}$Department of Computer Science,
  The University of Sheffield, Sheffield, UK \\
$^{4}$Department of Automatic Control Systems Engineering,
  The University of Sheffield, Sheffield, UK \\
$^{5}$Department of Computer Science,
  The University of Patras, Patras, Greece \\
}
% The Corresponding Author should be marked with an asterisk. Provide
% the exact contact address (this time including street name and city
% zip code) and email of the corresponding author
\def\corrAuthor{Kevin Gurney}
\def\corrAddress{Department of Psychology, The University of Sheffield,
  Western Bank, Sheffield, S10 2TP, UK}
\def\corrEmail{k.gurney@sheffield.ac.uk}

@@

@@

To the \ccg~model, we have added a rate-coded brainstem model (a ``saccadic
burst generator''), and a biomechanical eye, implemented using the
biomechanical modelling framework OpenSim. These will be described below,
but first we will give a description of the co-ordinate systems and the
modelling framework used.

\subsubsection{Coordinates in the world}

Before describing the biomechanical eye and the brain model, which consisted
of retinotopically mapped neural sheets, we describe the coordinate system
used in the world. The eye was located at the origin of a 3 dimensional, right-handed 
Cartesian co-ordinate system, with its fovea directed in the $-z$ direction. 
There was a notional spherical screen which was also centered at the origin of the
co-ordinate system and had a radius of 50 (in arbitrary units). The \e{fixation 
point} was the point on the
screen at which the eye was initially directed.
Onto the screen were projected luminances, each of which having a position 
described by two co-ordinates; $\theta_x$, a 
rotation of the horizon plane about the $x$ axis, and $\theta_y$, a rotation 
of the meridian plane about the $y$ axis. The position is the intersection
of these rotated planes with the spherical screen (disregarding
the intersection point of these three surfaces behind the eye).
%
Note that a luminance with positive $\theta_x$ was above the horizon of this world; 
one whose $\theta_y$ was positive lay to the left of the world's meridian.

Luminances were crosses of height and width subtending 6$\degree$ and whose
`bars' were 2$\degree$ thick. Luminances were oriented like $+$ symbols with
their vertical bar aligned with the meridian plane and their horizontal bar 
aligned with the horizon plane described above.

The eye's frame of reference was initially aligned with the world's frame
of reference. 
At each timestep, the eye's rotational state ($R_x$, $R_y$, $R_z$) was 
used to translate the three dimensional
Cartesian co-ordinates of the luminances in the world frame into co-ordinates
in the eye frame. The luminance coordinates in the eye's frame of reference
were used to determine the input to the brain model.

@@

Each population of 2500 neural elements was arranged in a 50 by 50
grid, with positions on the grid representing a retinotopic
mapping similar to that found empirically both in the superior 
colliculus (\cite{ottes_visuomotor_1986}) and in visual cortex (\cite{eric_l._schwartz_computational_1980}) and assumed in
this work to persist throughout the oculomotor system.

In a retinotopic mapping, the Cartesian coordinates of the light-sensitive
cells in the retina, whose density varies with distance from the fovea,
are transformed into the Cartesian co-ordinates of 
the correspondingly active cells on the colliculus. The mapping ensures 
that an even density of cells can be maintained in the colliculus, but 
ensures that a single group of proximal, active, retinal neurons will 
always activate a single, proximal group of neurons on the collicular 
surface.

The mapping turns out to resemble polar coordinates. That is, one 
axis of the collicular surface specifies the eccentricity of
a retinal location (how far it is from the fovea) and the second axis
specifies the rotational angle of the retinal location; we therefore
use the convention 
of referring to the eccentricity axis on the colliculus as $r$ and the 
rotation axis as $\phi$.

The \e{cortical magnification factor}, $M(r)$, gives the relationship
between the
radial eccentricity $r$ and the retinal neural density. As in 
\cite{cope_basal_2017}, we use a first-order approximation of the form
for $M(r)$ given in  \cite{rovamo_estimation_1979}:
\begin{equation} \label{eq:cmf}
M(r) = \frac{M_f}{1+\frac{r}{E_2}}
\end{equation}
The foveal magnification, $M_f$, is the magnification of the most central
region of the retina and has a value in the human of about
7.8 mm/$\degree$~\cite{rovamo_estimation_1979}.

In our model, $M_f$ is related to $W_{nfs}$, the width of the retinotopic neural 
field, $W_{fov}$, the width of the eye's field of view and $E_2$, the 
eccentricity at which the retinal density has halved by:
\begin{equation} \label{eq:fm}
   M_f = \frac {W_{nfs}} {E_2\;\ln\left(\frac{W_{fov}}{2 E_2} + 1\right)}
\end{equation}
Here, $W_{nfs}$ is 50, $W_{fov}$ is set to 61$\degree$, a
reduction from the biophysically accurate 150$\degree$ due to the small
number of neurons in the retinotopic neural field. $E_2$ is
2.5~\cite{cope_basal_2017,slotnick_electrophysiological_2001}.

The mapping from the retinotopic co-ordinates $r$ and
$\phi$ to rotational co-ordinates $\theta_x$  and $\theta_y$ on
the retina was written
down by \cite{eric_l._schwartz_computational_1980} for striate cortex and 
\cite{ottes_visuomotor_1986} for superior colliculus. We used the following
statement of this mapping for all of the populations in
the brain model:
\begin{equation}
   \theta_x = E_2 \left(e^{\frac{r}{M_f E_2}} - 1\right).\cos\left(\frac{2 \pi \phi}{W_{nfs}}\right)
\end{equation}
\begin{equation}
   \theta_y = E_2 \left(e^{\frac{r}{M_f E_2}} - 1\right).\sin\left(\frac{2 \pi \phi}{W_{nfs}}\right)
\end{equation}

Note that the mapping encompases the entire visual field. The value of
$\phi$ is allowed to vary from 0$\degree$ to 360$\degree$ along its axis.
Effectively, the two contralateral colliculi found in the biology are 
incorporated into a
single, square map, avoiding the need to carry out the kind of `colliculus
gluing' described in \cite{tabareau_geometry_2007}.

It is straightforward to show that the reverse mapping is given by:
\begin{equation}
   \phi = \frac{W_{nfs}} {2 \pi}\arctan\left(\frac{\theta_y}{\theta_x}\right)
\end{equation}
\begin{equation}
   r = M_fE_2\,\ln\left(\frac{1}{E_2}\sqrt{{\theta_x}^2 + {\theta_y}^2}+1\right)
\end{equation}

Figure \ref{fig:mapping} shows the result of the mapping for a view of
two cross-shaped luminances. One cross illuminates the fovea, which
results in a large comb-shape of activity. The more peripheral cross
produces (in FEF) an indistinct object centered at a larger value of
$r$.

@@

In this model, we implement a saccadic burst generator (SBG) based
on \cite{gancarz_neural_1998}. The SBG network for two of the model's
six channels is shown in Fig.~\ref{fig:sbg}. Activity from the output
layer of superior colliculus (SC\_avg) is fed into each channel, which
sums the activity it receives and processes it in populations each of
a single neural element representing all the neurons in that
population.

Neurons in the deep
layers of the superior colliculus called long lead burst neurons
innervate excitatory burst neurons in the reticular formation (REF,
and check if LLBNs are really in SC).

Alex B to write this section?

Note that inhibition from IBN populations is fed back - or is this
detail for the `integration' subsection?

@@

% Possibly introductory paragraph here.

% Patras - what about a figure here?

@@

%EOM dynamics
The force applied by EOMs is controlled by an excitatory signal
supplied by motoneurons in the brainstem. The dynamics of muscular
forces can be split into: 1) The elasticity of the muscles. 2) A delay
between the onset of the afferent excitatory signal and the actual
muscle contraction, caused by the transmission time of the action
potentials and by the necessary calcium release at the muscle
fibres. We developed a custom extraocular muscle model which captures
these features.

@@

% Brain to SBG, seeing as that's how I explained it above.
A number of studies have considered the form of the connection between
the deeper layers of the superior colliculus and the saccadic burst
generator~\cite{van_gisbergen_experimental_1985,ottes_visuomotor_1986,waitzman_superior_1991,groh_converting_2001,arai_two-dimensional_1994,goossens_dynamic_2006,tabareau_geometry_2007,van_opstal_linear_2008,goossens_optimal_2012},
which has become known as the spatial temporal transform (STT).  The
spatial aspect of the transform is thought to be implemented by a
weight-mapping~\cite{tabareau_geometry_2007,arai_two-dimensional_1994} and we
follow this idea.
% Words along the lines of:
Arai and co-workers trained a 20x20 neural network model of the
superior colliculs to discover the weight map under the assumption of
2D Gaussian activation profiles~\cite{arai_two-dimensional_1994}.
%
The training approach of Arai~et.~al. was not feasible in this study
due to the length of time required to run our model and its
stochasticity, which means multiple runs of the model were necessary
in order to generate output statistics.
%
Tabareau and co-workers wrote down a theoretical form of the weight
map, which follows from the mapping of \cite{ottes_visuomotor_1986}
and the assumption of invariant 2D Gaussian activity profiles in
SC~\cite{tabareau_geometry_2007}. As they found it closely resembles the
results of Arai et al., and it is a simple formulation, we considered
it as the means to generate the six weight maps in our own model.
%
One barrier to the use of the weight map
in \cite{tabareau_geometry_2007} was the \ccg~model's violation of the
\e{invariant integral hypothesis}. This states that the number of spikes
emitted by a neural element during a saccade (or in our model, the
integral of the neuron's output during the saccade) should be a
function only of its position within the hill of collicular
activity. That is, for any time-dependent hill of activity
$\mathcal{A}(\mathbf{z},t)$ at $\mathbf{z} = (r,\phi)$ on the collicular
surface, the integrated activity  $A_{\mathbf{x}}$ in an element at a
vector $\mathbf{x}$ away from $\mathbf{z}$ is
\begin{equation}
A_{\mathbf{x}} = \int_t \mathcal{A}(\mathbf{z}-\mathbf{x}, t) dt
\end{equation}
which is invariant for all $\mathbf{z}$. However, the very mapping on
which the \cite{tabareau_geometry_2007} result is based leads to a
very \e{variant} activity profile in the \ccg~model. A luminance of
a given size which excites activity near to the fovea causes activity
in a large number of neurons, whereas activity far from the fovea
excites a much smaller region. This effect is clearly demonstrated in
Fig.~\ref{fig:mapping} for equal sized targets both on and distal from
the fovea.

This led us to hypothesize that the retinotopic mapping be accompanied
by an associated widening projection field such that the hill of
activity in superior colliculus is invariant with position on the
collicular surface. There are a number of locations in the system in
which this widening projection field could exist. It could be
implemented in the projections between the retinal populations and the
superficial layer of SC along with the projection between the World
and the FEF population. However, this would affect activity within the
basal ganglia of the model, contradicting a result
in \cite{cope_basal_2017} which explains the `hockey stick' profile
for saccade latency as a function of saccade eccentricity. Instead, we
suggest that a widening projection field is encoded within the
superior colliculus itself, a complex, multi-layered structure which
could quite plausibly support such a function. Indeed, such widening
activity can be seen in the stimulation experiments
in \cite{vokoun_intralaminar_2010}
and \cite{vokoun_response_2014}. Although in this work we do not model
the SC in detail, we extended the model with a third functional layer
(\ccg has two layers; SC\_sup) and SC\_deep) named SC\_deep2. We
introduced a widening projection based on a Gaussian projection field
whose width, $\sigma(r)$ varies in inverse proportion to the 
magnification factor, $M(r)$, given in Eq.~\ref{eq:cmf} according to:
\begin{equation} \label{eq:sigmar}
\sigma(r) = \frac{m_{\sigma}}{M(r)} - \frac{m_{\sigma}}{M^0} + \sigma_0 \qquad r > r_0
\end{equation}
$m_{\sigma}$ is a scalar parameter which determines the `magnitude
of the widening'. $M^0$ is the `starting' magnification factor; 
within the foveal region ($0 \leq r \leq r_0$), the projection field 
is not allowed to widen and so
\begin{equation} \label{eq:sigmar2}
\sigma(r) = \sigma_0 \qquad r \leq r_0
\end{equation}
which makes $\sigma_0$ the width of the Gaussian projection field
within the foveal region. Note that the width of the foveal region 
for this projection is not identical to the foveal width used
in \_\_\_\_ \cmnt{Where else?}.
The \e{Widening Gaussian} projection weight, $w(r,d)$ is then 
computed as:
\begin{equation} \label{eq:widening}
w(r,d) = e^{-\frac{d^2}{2\sigma\left(r\right)^2}}
\end{equation}
where $d$ is the distance between the source and destination 
elements in the collicular plane.  $m_\sigma$ was set to 50, $\sigma_0$
was 0.3, $M^0$ was 12.43 and $r_0$ was 20.
% Could have a figure here, made from widening_gaussian.m

A further issue regarding the use of the theoretical weight map
in \cite{tabareau_geometry_2007} was that it does not consider the 
existence of the oblique extraocular muscles. There is indeed evidence that
only two dimensional information is encoded in superior
colliculus~\cite{wurtz_activity_1972,van_opstal_two-rather_1991}, but the
eye is actuated by six extraocular muscles. In order to find out a 
possible form for the input to the oblique muscles we carried out a 
training process described in supplementary material which depended on a
centroid computation in SC\_deep. For the four rectus muscles, the 
resulting weight maps resembled those found by Arai~et~al. The trained 
maps for the oblique muscles had a form very close to those for the 
inferior and superior rectus channels, but with a smaller magnitude.
The inferior oblique map resembled the superior rectus map and the 
superior oblique map resembled the inferior rectus. When parameterising 
the theoretical weight maps, we set the inferior/superior 
oblique maps to be 1/10$^{\mathrm{th}}$ of the superior/inferior rectus
maps, respectively. Interestingly, this suggests that there is a built-in
synergy between the vertical and oblique channels in the eye, although
the results will show there is some systematic change in the oblique error
with saccade end-point location.

@@

@@

@@

% Be sure to say that ``in future work, it will be possible to
% investigate saccade trajectory modifications in Parkinson's''

>>>>

% The Frontiers template.
\documentclass{frontiersSCNS}

@@

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{James {et~al.}}
\def\Authors{Sebastian James\,$^{1,2,*}$, Chris Papapavlou\,$^5$, Alexander Blenkinsop\,$^{1,2}$, \\
  Alexander Cope\,$^3$,  \\
  Sean Anderson\,$^{2,4}$, Konstantinos Moustakas\,$^5$, Kevin Gurney\,$^{1,2}$}
% Affiliations should be keyed to the author's name with superscript
% numbers and be listed as follows: Laboratory, Institute, Department,
% Organization, City, State abbreviation (USA, Canada, Australia), and
% Country (without detailed address information such as city zip codes
% or street names).
%
% If one of the authors has a change of address, list the new address
% below the correspondence details using a superscript symbol and use
% the same symbol to indicate the author in the author list.
\def\Address{\\
$^{1}$Adaptive Behaviour Research Group, Department of Psychology,
  The University of Sheffield, Sheffield, UK \\
$^{2}$Insigneo Institute for in-silico Medicine,
  The University of Sheffield, Sheffield, UK \\
$^{3}$Department of Computer Science,
  The University of Sheffield, Sheffield, UK \\
$^{4}$Department of Automatic Control Systems Engineering,
  The University of Sheffield, Sheffield, UK \\
$^{5}$Department of Electrical and Computer Engineering,
  The University of Patras, Patras, Greece \\
}
% The Corresponding Author should be marked with an asterisk. Provide
% the exact contact address (this time including street name and city
% zip code) and email of the corresponding author
\def\corrAuthor{Kevin Gurney}
\def\corrAddress{Department of Psychology, The University of Sheffield,
  Western Bank, Sheffield, S10 2TP, UK}
\def\corrEmail{k.gurney@sheffield.ac.uk}

@@

\cmnt{Background, gap, solution (cheers Sean!)}

@@

To the \ccg~model, we have added a rate-coded brainstem model (a `saccadic
burst generator'), and a biomechanical eye, implemented using the
biomechanical modelling framework OpenSim. These will be described below,
but first we will give a description of the co-ordinate systems and the
modelling framework used.

\subsubsection{Co-ordinates in the world}

Before describing the biomechanical eye and the brain model, which consisted
of retinotopically mapped neural sheets, we describe the co-ordinate system
used in the world. The eye was located at the origin of a 3 dimensional, right-handed
Cartesian co-ordinate system, with its fovea directed in the $-z$ direction.
There was a notional spherical screen which was also centered at the origin of the
co-ordinate system and had a radius of 50 (in arbitrary units). The \e{fixation
point} was the point on the
screen at which the eye was initially directed.
Onto the screen were projected luminances, each of which having a position
described by two co-ordinates; $\theta_{x}^{t}$, a
rotation of the horizon plane about the $x$ axis, and $\theta_{y}^{t}$, a rotation
of the meridian plane about the $y$ axis. The position is the intersection
of these rotated planes with the spherical screen (disregarding
the intersection point of these three surfaces behind the eye).
%
Note that a luminance with positive $\theta_{x}^{t}$ was above the horizon of this world;
one whose $\theta_{y}^{t}$ was positive lay to the left of the world's meridian. For this
reason, many of the figures in this paper are plotted with $-\theta_{y}$ on the
$x$-axis and $\theta_{x}$ on the $y$-axis so that targets that lay up and to the right
in the world do so in the graphs, also.

Luminances were crosses of height and width subtending $\pm$3\dg~and whose
`bars' were 2\dg~thick. Luminances were oriented like $+$ symbols with
their vertical bar aligned with the meridian plane and their horizontal bar
aligned with the horizon plane described above.

The eye's frame of reference was initially aligned with the world's frame
of reference.
At each timestep, the eye's rotational state (described by the Euler
rotations $\theta_x$, $\theta_y$, $\theta_z$) was
used to translate the three dimensional
Cartesian co-ordinates of the luminances in the world frame into co-ordinates
in the eye frame. The luminance co-ordinates in the eye's frame of reference
were used to determine the input to the brain model.

@@

Each population of 2500 neural elements was arranged in a 50 by 50
grid, with positions on the grid representing a retinotopic
mapping similar to that found empirically both in the superior
colliculus [\cite{ottes_visuomotor_1986}] and in visual cortex
[\cite{eric_l._schwartz_computational_1980}] and assumed in
this work to persist throughout the oculomotor system.

In a retinotopic mapping, the Cartesian co-ordinates of the light-sensitive
cells in the retina, whose density varies with distance from the fovea,
are transformed into the Cartesian co-ordinates of
the correspondingly active cells on the colliculus. The mapping ensures
that an even density of cells can be maintained in the colliculus, but
ensures that a single group of proximal, active, retinal neurons will
always activate a single, proximal group of neurons on the collicular
surface.

The mapping turns out to resemble polar co-ordinates. That is, one
axis of the collicular surface specifies the eccentricity of
a retinal location (how far it is from the fovea) and the second axis
specifies the rotational angle of the retinal location; we therefore
use the convention of referring to the eccentricity axis on the
colliculus as $r$ and the rotation axis as $\phi$.

The \e{cortical magnification factor}, $M(r)$, gives the relationship
between the
radial eccentricity $r$ and the retinal neural density. As in
\cite{cope_basal_2017}, we use a first-order approximation of the form
for $M(r)$ given in  \cite{rovamo_estimation_1979}:
\begin{equation} \label{eq:cmf}
M(r) = \frac{M_f}{1+\frac{r}{E_2}}
\end{equation}
The foveal magnification, $M_f$, is the magnification of the most central
region of the retina and has a value in the human of about
7.8 mm/\dg~\cite{rovamo_estimation_1979}.

In our model, $M_f$ is related to $W_{nfs}$, the width of the retinotopic neural
field, $W_{fov}$, the width of the eye's field of view and $E_2$, the
eccentricity at which the retinal density has halved by:
\begin{equation} \label{eq:fm}
   M_f = \frac {W_{nfs}} {E_2\;\ln\left(\frac{W_{fov}}{2 E_2} + 1\right)}
\end{equation}
Here, $W_{nfs}$ is 50 (the side length of the 50x50 grid)
and $W_{fov}$ is set to 61\dg, a
reduction from the biophysically accurate 150\dg~due to the small
number of neurons in the retinotopic neural field. $E_2$ is
2.5 [\cite{cope_basal_2017,slotnick_electrophysiological_2001}].

The mapping from the retinotopic co-ordinates in the brain to rotational
co-ordinates of the stimulus/response was written down by
\cite{schwartz_spatial_1977,eric_l._schwartz_computational_1980}
for measurements of
striate cortex [visual stimulus to electrophysiological response---\cite{daniel_representation_1961, talbot_physiological_1941}]
and by \cite{ottes_visuomotor_1986} for superior colliculus data
[electrophysiological SC stimulus to eye movement
response---\cite{robinson_eye_1972}].
% Ottes et al. fit to stimulus-to-movement data, and give a
% movement-to-collicular-coordinate mapping in the paper.
We used the following statement of this mapping to introduce stimuli
into the `World' input population of the brain model:
\begin{equation}\label{eq:thetas_to_phi}
   \phi = \frac{W_{nfs}} {2 \pi}\arctan\left(\frac{\theta_{y}^{t}}{\theta_{x}^{t}}\right)
\end{equation}
\begin{equation}\label{eq:thetas_to_r}
   r = M_fE_2\,\ln\left(\frac{1}{E_2}\sqrt{{\theta_{x}^{t}}^2 + {\theta_{y}^{t}}^2}+1\right)
\end{equation}
Note that we use $r$ and $\phi$ as the co-ordinates on the `collicular
surface'. Schwartz uses $r$ and $\phi$ as the polar coordinates of the
retinal stimulus; Ottes et al.~use $r$ and $\phi$ as polar coordinates
for the eye movement response; both use $u$ and $v$ as the Cartesian co-ordinates of the neural map. We use $\theta_{x}^{t}$ and $\theta_{y}^{t}$
to give Euler rotations for the retinal target stimulus. Note also that
the form of Eqns.~\ref{eq:thetas_to_phi} \& \ref{eq:thetas_to_r} is slightly
different from that given in
\cite{ottes_visuomotor_1986} because our $\theta_{x}^{t}$ and
$\theta_{y}^{t}$ are not the polar co-ordinates used in that work.

Note that the mapping encompases the entire visual field. The value of
$\phi$ is allowed to vary from 0\dg~to 360\dg~along its axis.
Effectively, the two contralateral colliculi found in the biology are
incorporated into a
single, square map, avoiding the need to carry out the kind of `colliculus
gluing' described in \cite{tabareau_geometry_2007}.

It is straightforward to show that the reverse mapping is given by:
\begin{equation}
   \theta_x = E_2 \left(e^{\frac{r}{M_f E_2}} - 1\right).\cos\left(\frac{2 \pi \phi}{W_{nfs}}\right)
\end{equation}
\begin{equation}
   \theta_y = E_2 \left(e^{\frac{r}{M_f E_2}} - 1\right).\sin\left(\frac{2 \pi \phi}{W_{nfs}}\right)
\end{equation}
where we have dropped the $t$ superscript on $\theta_x$ \& $\theta_y$, as
these equations transform a collicular location into rotations of the eye.

Figure \ref{fig:mapping} shows the result of the mapping for a view of
two cross-shaped luminances. One cross illuminates the fovea, which
results in a large comb-shape of activity. The more peripheral cross
produces (in FEF) an indistinct object centred at a larger value of
$r$.

@@

We implemented a saccadic burst generator (SBG) based
on the connectivity outlined in \cite{gancarz_neural_1998}. The SBG network
for two of the model's
six channels is shown in Fig.~\ref{fig:sbg}. Activity from the output
layer of superior colliculus (SC\_avg) is fed into each channel, which
sums the activity it receives and processes it in populations each of
a single neural element representing all the neurons in that
population. Each channel of the SBG functions to create the motor neuron
activations that are required to accelerate the eye in a particular
direction, then hold the eye in its new position against the returning
force generated by the elastic properties of the muscles.  The
required motor neuron activations are therefore a brief burst of
increased activity that accelerates the eye, followed by a period
of activity that is less than the burst firing rate but higher than
the tonic rate that exists when the eye is at the centre.  This
holds the eye in its new position.

The SBG connectivity produce each of the these elements separately,
then sums them to create the desired `bump and tonic' activation
time series.  The input to the first population in the SBG, the
long-lead burst neurons (LLBNs), is conceived as originating from
one of the deep layers of the superior colliculus.  The activity
of the LLBNs are passed to excitatory burst neurons (EBNs) which,
in turn, inhibit the LLBNs via the activity of the inhibitory
burst neurons (IBNs).  This feedback loop has a transmission delay,
which allows activity to build up in the EBNs before the inhibition
is activated and the activity is then reduced again.  This mechanism
generates the `bump'.

The generation of the `tonic' phase of the required time series is
achieved simply by integrating the bump over time and multiplying by a
some small gain factor.  This is the function of the tonic neurons (TNs).
The firing rate of the motor neuron defines the amount of force applied
to the eye by that muscle.  Thus, the size of the `bump' defines how far
the eye moves in that channel's direction.  The gain and delay parameters
in the LLBN-EBN-IBN-LLBN feedback loop therefore have to be tuned such that
the endpoint of the saccade is reasonably accurate.  Furthermore the
restoring force generated by the elasticity of the muscles is dependent
on the radial distance.  The value of the new tonic firing rate, after
the `bump' is dependent on the end location of the eye.  If the ratio
between the EBN firing rate and the TN firing rate is not exactly
correct, the eye will drift away from the saccade endpoint after the
saccade has been completed.  The EBN-TN connection strength is
therefore tuned such that the TN firing rate yields a stable eye
position across a range of eye eccentricities.

The omnipause neurons (OPNs) are tonically active and inhibit the EBNs.
The activity of the OPNs is itself inhibited by activity in the LLBNs.
The purpose of this arrangement is to ensure the eye does not move in
response to neural noise.

Each mean activity of all the neurons in each SBG population (except the
TNs) is defined by a single leaky integrator, first order differential equation.
\begin{equation}\label{eq:thetas_to_r}
   \frac{da}{dt} = \frac{1}{\tau}(y-a)
\end{equation}
where $a$ is the activation of the nucleus, and $\tau$ is the time constant
of the nucleus. $y$ is a piecewise linear function of the weighted sum of
inputs to the nucleus and is given by
\begin{equation}
       y(IN) = \begin{cases}
       0    & IN \leq b \\
      IN-b   & b \leq IN \leq 1+b \\
      1   & IN \geq 1+b \\
   \end{cases}
\end{equation}
where $b$ is the $IN$ axis offset.  $IN$ is the weighted sum of inputs to
the nucleus and is given by,
\begin{equation}
    IN = \sum_{m}^{M} w_{mn} a_{m}
\end{equation}
where $a_{m}$ is the activation of the $m^{th}$ afferent nucleus.
$w_{mn}$ is the connection strength between the $m^{th}$ afferent nucleus
and the current nucleus. The activity of the TNs are defined as
\begin{equation}
   \frac{da}{dt} = \frac{1}{\tau}\,y
\end{equation}
with an identical piecewise linear transfer function as the other SBG populations.

\cmnt{Note that inhibition from IBN populations is fed back - or is this
detail for the `integration' subsection?}

@@

% Possibly introductory paragraph here.
The output signals of the brainstem are used to drive the biomechanical model. The latter is not only used to get tangible feedback on the simulated saccades including motion trajectories, but adds one more modeling dimension related to the inertial properties of the eye plant including muscle properties.

% Patras - what about a figure here? --- WILL ADD TOMORROW TUESDAY

@@

%EOM dynamics
The force applied by EOMs is controlled by an excitatory signal
supplied by motoneurons in the brainstem. The neural drive to produce a saccadic eye movement can be characterized by a pulse component to overcome the viscoelasticity of the orbital plant, a step component to stabilize the eye in the new position, and a slide component that models the gradual transition between the pulse and step.

The dynamics of muscular
forces can be split into: 1) The elasticity of the muscles. 2) A delay
between the onset of the afferent excitatory signal and the actual
muscle contraction, caused by the transmission time of the action
potentials and by the necessary calcium release at the muscle
fibres. We developed a custom extraocular muscle model which captures
these features.

@@

% Brain to SBG, seeing as that's how I explained it above.
A number of studies have considered the form of the connection between
the deeper layers of the superior colliculus and the saccadic burst
generator~\cite{van_gisbergen_experimental_1985,ottes_visuomotor_1986,waitzman_superior_1991,groh_converting_2001,arai_two-dimensional_1994,goossens_dynamic_2006,tabareau_geometry_2007,van_opstal_linear_2008,goossens_optimal_2012},
which has become known as the spatial temporal transform (STT).  The
spatial aspect of the transform is thought to be implemented by a
weight-mapping~\cite{tabareau_geometry_2007,arai_two-dimensional_1994} and we
follow this idea.
% Words along the lines of:
Arai and co-workers trained a 20x20 neural network model of the
superior colliculs to discover the weight map under the assumption of
2D Gaussian activation profiles~\cite{arai_two-dimensional_1994}.
%
The training approach of Arai~et.~al. was not feasible in this study
due to the length of time required to run our model and its
stochasticity, which means multiple runs of the model were necessary
in order to generate output statistics.
%
Tabareau and co-workers wrote down a theoretical form of the weight
map, which follows from the mapping of \cite{ottes_visuomotor_1986}
and the assumption of invariant 2D Gaussian activity profiles in
SC~\cite{tabareau_geometry_2007}. As they found it closely resembles the
results of Arai et al., and it is a simple formulation, we considered
it as the means to generate the six weight maps in our own model.
%
One barrier to the use of the weight map
in \cite{tabareau_geometry_2007} was the \ccg~model's violation of the
\e{invariant integral hypothesis}. This states that the number of spikes
emitted by a neural element during a saccade (or in our model, the
integral of the neuron's output during the saccade) should be a
function only of its position within the hill of collicular
activity. That is, for any time-dependent hill of activity
$\mathcal{A}(\mathbf{z},t)$ at $\mathbf{z} = (r,\phi)$ on the collicular
surface, the integrated activity  $A_{\mathbf{x}}$ in an element at a
vector $\mathbf{x}$ away from $\mathbf{z}$ is
\begin{equation}
A_{\mathbf{x}} = \int_t \mathcal{A}(\mathbf{z}-\mathbf{x}, t)\;dt
\end{equation}
which is invariant for all $\mathbf{z}$. However, the very mapping on
which the \cite{tabareau_geometry_2007} result is based leads to a
very \e{variant} activity profile in the \ccg~model. A luminance of
a given size which excites activity near to the fovea causes activity
in a large number of neurons, whereas activity far from the fovea
excites a much smaller region. This effect is clearly demonstrated in
Fig.~\ref{fig:mapping} for equal sized targets both on and distal from
the fovea.

This led us to hypothesize that the retinotopic mapping be accompanied
by an associated widening projection field such that the hill of
activity in superior colliculus is invariant with position on the
collicular surface. There are a number of locations in the system in
which this widening projection field could exist. It could be
implemented in the projections between the retinal populations and the
superficial layer of SC along with the projection between the World
and the FEF population. However, this would affect activity within the
basal ganglia of the model, contradicting a result
in \cite{cope_basal_2017} which explains the `hockey stick' profile
for saccade latency as a function of saccade eccentricity. Instead, we
suggest that a widening projection field is encoded within the
superior colliculus itself, a complex, multi-layered structure which
could quite plausibly support such a function. Indeed, such widening
activity can be seen in the stimulation experiments
in \cite{vokoun_intralaminar_2010}
and \cite{vokoun_response_2014}. Although in this work we do not model
the SC in detail, we extended the model with a third functional layer named SC\_deep2
(\ccg~has only the two layers SC\_sup and SC\_deep). We
introduced a widening projection based on a Gaussian projection field
whose width, $\sigma(r)$ varies in inverse proportion to the
magnification factor, $M(r)$, given in Eq.~\ref{eq:cmf} according to:
\begin{equation} \label{eq:sigmar}
\sigma(r) = \frac{m_{\sigma}}{M(r)} - \frac{m_{\sigma}}{M^0} + \sigma_0 \qquad r > r_0
\end{equation}
$m_{\sigma}$ is a scalar parameter which determines the `magnitude
of the widening'. $M^0$ is the `starting' magnification factor;
within the foveal region ($0 \leq r \leq r_0$), the projection field
is not allowed to widen and so
\begin{equation} \label{eq:sigmar2}
\sigma(r) = \sigma_0 \qquad r \leq r_0
\end{equation}
which makes $\sigma_0$ the width of the Gaussian projection field
within the foveal region. (Note that the value chosen for the width
of the foveal region, $r_0$ is not identical to the foveal shift parameter used
in the \e{DecayingAtFovea} projections into striatum.)
The \e{Widening Gaussian} projection weight, $w(r,d)$ is then
computed as:
\begin{equation} \label{eq:widening}
w(r,d) = e^{-\frac{d^2}{2\sigma\left(r\right)^2}}
\end{equation}
where $d$ is the distance between the source and destination
elements in the collicular plane.  $m_\sigma$ was set to 50, $\sigma_0$
was 0.3, $M^0$ was 12.43 and $r_0$ was 20.
% Could have a figure here, made from widening_gaussian.m

A further issue regarding the use of the theoretical weight map
in \cite{tabareau_geometry_2007} was that it does not consider the
existence of the oblique extraocular muscles. There is indeed evidence that
only two dimensional information is encoded in superior
colliculus~\cite{wurtz_activity_1972,van_opstal_two-rather_1991}, but the
eye is actuated by six extraocular muscles. In order to find out a
possible form for the input to the oblique muscles we carried out a
training process described in supplementary material which depended on a
centroid computation in SC\_deep. For the four rectus muscles, the
resulting weight maps resembled those found by
\cite{arai_two-dimensional_1994}. The trained
maps for the oblique muscles had a form very close to those for the
inferior and superior rectus channels, but with a smaller magnitude.
The inferior oblique map resembled the superior rectus map and the
superior oblique map resembled the inferior rectus. When parameterising
the theoretical weight maps, we set the inferior/superior
oblique maps to be 1/10$^{\mathrm{th}}$ of the superior/inferior rectus
maps, respectively. Interestingly, this suggests that there is a built-in
synergy between the vertical and oblique channels in the eye, although
the results will show there is some systematic change in the oblique error
with saccade end-point location.

@@

The final models, on which the results of this paper are based are
named `TModel3', `TModel4' and `TModel5'. Descriptions of earlier
versions of the model can be
found in the code repository given in SUPPLEMENTAL DATA.

@@

We now present results derived from the fully parameterised and integrated
model; where we took advantage of the fact that it is a closed loop system. This
allowed us to present sequences of target luminances and allow the model to
direct its fovea at the most salient target.

@@

The aim of this study was to demonstrate the importance of modelling
neurological systems \e{in concert with} the biomechanical systems
together with which they have evolved. We hypothesised that by combining
existing neurophysiological models with an accurate
model of a musculo-skeletal system, and then `closing the loop' by allowing
the movements of the virtual muscles to modulate sensory feedback to
the brain model, shortcomings in the constituent models would be
revealed, leading to new knowledge. To demonstrate the validity
of this hypothesis, we built an integrated model and then identified the
modifications which were necessary to give it the ability to make
accurate movements under one type of stimulus. We then
examined its behaviour with other stimuli.
